### 2021-04-09 Submission template.Rmd
Submission template

---
header-includes:
   - \usepackage{bm}
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \usepackage{booktabs}
   - \usepackage{sectsty} \sectionfont{\centering \emph}
output:
  pdf_document: default
  html_document: default
  word_document: default

---
\newcommand\btheta{\mbox{\boldmath${\theta}$}}

```{r setup, include=FALSE}
# Use echo = FALSE for Answer Key without R code, 
# echo = TRUE for complete solutions.
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library("tidyverse")
library("xtable")
library("Rcpp")
library("plyr")
library("knitr")
```

Team No: 202

Track No.: 2.

# Our name of the project

Link to GitHub Repository: https://github.com/MantautasRimkus/StanfordDatathonCSU

### Introduction

### 2021-04-10 Submission template.Rmd
Template updated with a plan

@@
 
-### Introduction
+### Plan
+
+* Write introduction: About Covid, about anxiety, about how 19-28 more(no more?) prone to it. Write that our analysis shows that a lot of 18-29 are in the college.
+
+* Write introduction: calculate (using historical data) how many 18-29 is in college. Maybe prediction for 2020. 
+* Describe data sources used in this research:
+   - Anxiety
+   - Covid
+   - Google
+
+* Overview trends in anxiety data (functional analysis)
+
+* Panel modelling
+
+* Discussion/results.

### 2021-04-10 plot-covid.R
Added code to aggregate covid data to the necessary form (Weekly, Log Ratios)

@@ -1,0 +1,59 @@
+library(lubridate)
+library(dplyr)
+library(ggplot2)
+library(zoo)
+
+# Read covid data
+covid <- read.csv("covid-data/United_States_COVID-19_Cases_and_Deaths_by_State_over_Time.csv")
+
+states <- c("AK", "AL", "AR", "AZ", "CA", "CO", "CT", "DE", "FL", "GA",
+            "HI", "IA", "ID", "IL", "IN", "KS", "KY", "LA", "MA", "MD",
+            "ME", "MI", "MN", "MO", "MS", "MT", "NC", "ND", "NE", "NH",
+            "NJ", "NM", "NV", "NY", "OH", "OK", "OR", "PA", "RI", "SC",
+            "SD", "TN", "TX", "UT", "VA", "VT", "WA", "WI", "WV", "WY",
+            "NYC") # Include NYC, to be combined with NY
+
+# Process dates for sorting
+covid <- data.frame(
+  Date = as.Date(strptime(covid$submission_date, "%m/%d/%Y")),
+  State = covid$state,
+  Cases = covid$new_case
+)
+covid <- covid[covid$State %in% states,]
+# Combine NYC with NY by taking first 2 substring
+covid$State <- substr(covid$State, 1, 2)
+covid <- aggregate(Cases ~ Date + State, covid, sum)
+# Add week
+covid$EndofWeek <- covid$Date - (wday(covid$Date) - 1) + 7 * (wday(covid$Date) != 1)
+
+
+# Aggregate to weekly and calculate sequential log ratios
+covid.weekly <- covid %>%
+  group_by(State, EndofWeek) %>%
+  summarize(Cases = mean(Cases))
+covid.weekly <- covid.weekly %>%
+  arrange(State, EndofWeek) %>%
+  group_by(State) %>%
+  mutate(LogRatio = log(Cases / lag(Cases)))
+covid.weekly$LogRatio <- replace(covid.weekly$LogRatio, is.infinite(covid.weekly$LogRatio), NaN)
+  
+# Plot each state as a series
+ggplot(covid.weekly, aes(x=EndofWeek, y=LogRatio)) + geom_line(aes(color=State))
+
+#############################################################################
+### Aggregate to the US
+
+# Aggregate
+covid.us <- aggregate(Cases ~ Date + EndofWeek, covid, sum)
+
[…]

### 2021-04-10 submissiondraftNR.Rmd
draft of introduction

title: 'Draft of Report for Datathon'
author: "Nathan Ryder Mantautas Rimkus, Paul Chong, Ian Taylor"
date: '`r format(Sys.time(), "%d %B %Y")`'
output:
  bookdown::html_document2: 
    number_sections: true
    toc: yes

header-includes:
  - \usepackage{amsmath}
  - \usepackage{enumerate}
  - \usepackage{booktabs}
---

# Introduction

University students have been found to experience mental health disorders at a disproportionate rate.  One meta-analysis found the average rate of depression in university students to be 30.6\%, when the general US population is estimated to have a 9\% average rate (REF).  The age range of most university students, 18-22, is host to large social and mental developments and is one of the more common time periods for mental disorders like anxiety to manifest (REF).  Adding the pressures of school and high-impact life choices, it is understandable that university students experience more than their fair share of mental health issues.

The year 2020 and the disease Covid-19 added further difficulty to the lives of those who are college aged.  Students who returned or were admitted to universities in fall 2020 were forced to learn new styles of learning and make sacrifices in their social lives.   Universities transitioned to partially or fully online curriculum, adding distance between student and staff communities.  Unsurprisingly, measures for anxiety and depression have increased under the duration of the Covid-19 pandemic.  This report examines the survey-reported changes in anxiety or depression prevalence for college-aged persons in the United States.  Data on Covid-19 infection rates are linked with the mental health surveys.

# Data

Figures:

plot of age groups anxiety

plot of covid?

### 2021-04-11 plot-covid.R
Code to save covid data csvs.

@@ -57,3 +57,11 @@ covid.us.weekly$LogRatio <- replace(covid.us.weekly$LogRatio, is.infinite(covid.
 
 
 ggplot(covid.us.weekly, aes(x=EndofWeek, y=LogRatio)) + geom_line()
+
+##############################################################################
+
+# output files after processing
+
+write.csv(covid.weekly, "covid-data/weekly-states.csv")
+
+write.csv(covid.us.weekly, "covid-data/weekly-us.csv")

### 2021-04-11 anxiety-covid-ts-model.R
Run time series and panel models on anxiety with covid as predictor

@@ -1,0 +1,97 @@
+# This file runs a time series model for anxiety of 18-29 year olds with
+# covid cases as a predictor
+
+library(plm)
+
+# Read in anxiety data, covid data (whole US) and covid data (by state)
+anxiety <- read.csv("anx_rate_ext.csv")
+covid.us <- read.csv("covid-data/weekly-us.csv")
+covid <- read.csv("covid-data/weekly-states.csv")
+
+# Join together data sources into data frame for linear model
+dates <- anx$date_end
+
+modeldata <- data.frame(
+  Week = dates,
+  anx = anxiety$US
+)
+for (i in 1:nrow(modeldata)) {
+  d <- modeldata[i, "Week"]
+  d.lag1 <- as.character(as.Date(d) - 7)
+  d.lag2 <- as.character(as.Date(d) - 14)
+  
+  # Anxiety from the previous week(s)
+  if (d.lag1 %in% anxiety$date_end) { 
+    modeldata[i, "anx.lag1"] <- anxiety[anxiety$date_end == d.lag1, "US"]
+  }
+  if (d.lag2 %in% anxiety$date_end) { 
+    modeldata[i, "anx.lag2"] <- anxiety[anxiety$date_end == d.lag2, "US"]
+  }
+  
+  # Covid cases (log ratio change) in current week and previous week(s)
+  if (d %in% covid.us$EndofWeek) {
+    modeldata[i, "covid"] <- covid.us[covid.us$EndofWeek == d, "LogRatio"]
+  }
+  if (d.lag1 %in% covid.us$EndofWeek) {
+    modeldata[i, "covid.lag1"] <- covid.us[covid.us$EndofWeek == d.lag1, "LogRatio"]
+  }
+  if (d.lag1 %in% covid.us$EndofWeek) {
+    modeldata[i, "covid.lag2"] <- covid.us[covid.us$EndofWeek == d.lag2, "LogRatio"]
+  }
+}
+modeldata$Election <- modeldata$Week %in% c("2020-11-01", "2020-11-08")
+
+
+################################################################################
+# Run linear models
+################################################################################
+
+# Model 1 - AR(1) of anxiety nationally, no covid
+fit.ar1anx <- lm(anx ~ anx.lag1, modeldata)
+summary(fit.ar1anx)
+
+# Model 2 - AR(2) of anxiety nationally, no covid
+fit.ar2anx <- lm(anx ~ anx.lag1 + anx.lag2, modeldata)
+summary(fit.ar2anx)
+
[…]

### 2021-04-11 submissiondraftIMT.Rmd
Added my submission draft. Ignored pdfs

--- 2021-04-10 Submission template.Rmd
@@
 * Discussion/results.
+
+## Predicting Anxiety from COVID Cases Using a Panel Model
+
+We are searching for a way for school administrators to anticipate increased anxiety among their students.
+In 2020 and today, anecdotally, COVID is a significant source of anxiety.
+We used a panel model to find the predictiveness of covid cases on anxiety in 18-29 year olds.
+
+Panel models are used for regression when the response and predictor variables are longitudinal.
+In this case, we treat US states as entities and have longitudinal values for both anxiety and new COVID cases.
+
+For states $s=1,\dots, 50$ and time periods $t=1,\dots,T$, we use the model
+
+$$
+\mbox{anx}_{s,t} = \alpha_s + \beta_1 \mbox{anx}_{s,t-1} + \beta_2 \mbox{COVID}_{s,t} + \varepsilon_{s,t}
+$$
+
+This model allows for state-specific intercepts ($\alpha_s$), autocorrelation in anxiety within a state ($\beta_1$), and an effect of COVID cases on anxiety ($\beta_1$).
+
+Alternatively, we can have a random effect for the state-specific intercepts,
+\begin{align*}
+\mbox{anx}_{s,t} &= \alpha + a_s + \beta_1 \mbox{anx}_{s,t-1} + \beta_2 \mbox{COVID}_{s,t} + \varepsilon_{s,t} \\
+a_s &\sim N(0, \sigma^2_a)
+\end{align*}
+
+The `plm` package in `R` allows us to fit both of these models and perform a hypothesis test to determine which fits our data better.

### 2021-04-11 Submission template_Mantautas.Rmd

---
header-includes:
   - \usepackage{bm}
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \usepackage{booktabs}
   - \usepackage{sectsty} \sectionfont{\centering \emph}
output:
  pdf_document: default
  html_document: default
  word_document: default

---
\newcommand\btheta{\mbox{\boldmath${\theta}$}}

```{r setup, include=FALSE}
# Use echo = FALSE for Answer Key without R code, 
# echo = TRUE for complete solutions.
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library("tidyverse")
library("xtable")
library("Rcpp")
library("plyr")
library("knitr")
library("gghighlight")
```

Team No: 202

Track No.: 2.

# Our name of the project

Link to GitHub Repository: https://github.com/MantautasRimkus/StanfordDatathonCSU

### Plan

#Data extrapolation

As described before, the data from Household Pulse Survey is named by weeks. As it is described in the technical documentation (LINK), these weeks do not represent the actual interval of 7 days. For example, "Week 1" refers to data that was collected between April 23, 2020, to May 5, 2020, (13 days interval), but Week 9 refers to data that was collected between June 25, 2020, to June 30, 2020 (6 days interval). Thus time points $s_i$ refers to the end date of each Week collection.

Notice that $s_i$ is sparse and leads to unequal spaces between data points. Thus to ensure univariate differences between time points (to be able to implement the models described later), we used interpolation, where our new time points $t_1,\ldots,t_{47}$ represents Sundays between May 05, 2020, to March 29, 2021. To calculated some data point $X_{t_i}$, where $s_{j-1}<t_i\leq s_{j}$:

$$
X_{t_i}=\frac{d(s_{j-1},t_i)}{d(s_{j-1},s_{j})}X_{s_j} + \frac{d(t_i,s_{j})}{d(s_{j-1},s_{j})}X_{s_{j-1}}
$$

An example of raw data $X_{s_1},\ldots,X_{s_27}$ and interpolated data $X_{t_1},\ldots, X_{t_j}$ interpolation of anxiety ratio for Colorado is given below.

```{r,fig.cap="Anxiety Ratio in Colorado. Black - original data points, Red = interpolated data with equal time spaces", }
data_raw <- read.csv("anx_Data_estimates.csv")
data_inter <- read.csv("anx_rate_ext.csv")

data_raw %>%
  dplyr::filter(State %in% c("CO")) %>%
  dplyr::filter(Age %in% c("18 - 29")) %>%
  dplyr::select(State,date_end,Response) %>%
  ggplot(aes(x=as.Date(date_end),y=Response)) +
  geom_line() +
  geom_point() +
  geom_line(data=data_inter[,c(2,8)],aes(x=as.Date(date_end),y=CO),col="red") +         geom_point(data=data_inter[,c(2,8)],aes(x=as.Date(date_end), y=CO),col="red") + 
   theme_bw() +
   ylab("Anxiety ratio") +
   xlab("Date")
```

The plot for anxiety ratio for all states with emphasis for Colorado and California

```{r}
data_inter[,-1] %>%
   tidyr::gather("State","Value",-date_end) %>%
   dplyr::filter(!State=="US") %>%
   ggplot(aes(x=as.Date(date_end),y=Value,col=State)) +
   geom_line() +
   gghighlight(State %in% c("CO","CA")) +
   theme_bw() +
   ylab("Anxiety ratio") +
   xlab("Date")

data_inter[,-1] %>%
   tidyr::gather("State","Value",-date_end) %>%
   dplyr::filter(!State=="US") %>%
   group_by(date_end) %>%
   dplyr::summarise(mean=mean(Value), sd=sd(Value)) %>%
   tidyr::gather("Statistics","Value",-date_end) %>%
   ggplot(aes(x=as.Date(date_end),y=Value,group=Statistics))+
   geom_line()+
   facet_wrap(~Statistics) +
   theme_bw()
```

```{r}
require(rmgarch)

````

### 2021-04-11 anxiety-covid-ts-model.R
Models with search trends

@@ -1,12 +1,24 @@
 # This file runs a time series model for anxiety of 18-29 year olds with
 # covid cases as a predictor
 
+dplyr
 library(plm)
 
 # Read in anxiety data, covid data (whole US) and covid data (by state)
 anxiety <- read.csv("anx_rate_ext.csv")
 covid.us <- read.csv("covid-data/weekly-us.csv")
 covid <- read.csv("covid-data/weekly-states.csv")
+search <- read.csv("statetrends.csv")
+
+########## Add diffs to search data
+`%>%` <- dplyr::`%>%`
+search <- search %>%
+  dplyr::arrange(location, date_end) %>%
+  dplyr::group_by(location) %>%
+  dplyr::mutate(
+    diff = hits - dplyr::lag(hits),
+    logratio = log(hits / dplyr::lag(hits))
+  )
 
 # Join together data sources into data frame for linear model
 dates <- anx$date_end
@@ -81,6 +93,20 @@ for (i in 1:nrow(paneldata)) {
   if (d %in% anxiety$date_end) {
     paneldata[i,"anx"] <- anxiety[anxiety$date_end == d, s]
   }
+  if (d %in% search$date_end) {
+    paneldata[i,"search"] <- search[
+      (search$State == s) & (search$date_end == d),
+      "hits"
+    ]
+    paneldata[i,"searchdiff"] <- search[
+      (search$State == s) & (search$date_end == d),
+      "diff"
+    ]
+    paneldata[i,"searchlogratio"] <- search[
+      (search$State == s) & (search$date_end == d),
+      "logratio"
+    ]
+  }
 }
 
 paneldata <- paneldata[paneldata$Date %in% dates,] # Only dates with an anxiety measurement
@@ -95,3 +121,15 @@ fit.states.random <- plm(anx ~ LogCasesRatio + lag(anx), paneldata,
 summary(fit.states.random)
 
 phtest(fit.states.fixed, fit.states.random)
+
+
+# Model: panel model by state with anxiety as response regressed on covid SEARCHES
+fit.states.search.fixed <- plm(anx ~ searchlogratio + lag(anx), paneldata, 
+                        index=c("State", "Date"), model="within")
+summary(fit.states.search.fixed)
+
+fit.states.search.random <- plm(anx ~ searchlogratio + lag(anx), paneldata, 
+                         index=c("State", "Date"), model="random")
+summary(fit.states.search.random)
+
+phtest(fit.states.search.fixed, fit.states.search.random)

### 2021-04-11 Anxiety_Data_Interpolation.R
Tidy R code for interpolation

@@ -1,0 +1,89 @@
+#############################################################################
+####This code for anxiety estimates interpolation
+####Output: File anx_rate_ext.csv
+###############
+
+############################################################################
+#Libraries 
+
+library(tidyverse)
+library(plyr)
+library(readr)
+library("lubridate")
+library(fda)
+
+###########################################################################
+#Setting up weeks dates. 
+
+dates_end<- c(Week1="May 5, 2020",Week2="May 12, 2020",Week3="May 19, 2020",
+              Week4="May 26, 2020",Week5="June 2, 2020",Week6="June 9, 2020",
+              Week7="June 16, 2020",Week8="June 23, 2020",Week9="June 30, 2020",
+              Week10="July 7, 2020",Week11="July 14, 2020",Week12="July 21, 2020",
+              Week13="August 31, 2020",Week14="September 14, 2020",
+              Week15="September 28, 2020",Week16="October 12, 2020",
+              Week17="October 26, 2020",Week18="November 9, 2020",
+              Week19="November 23, 2020",Week20="December 7, 2020",
+              Week21="December 21, 2020", Week22="January 18, 2021",
+              Week23="February 1, 2021",Week24="February 15, 2021",
+              Week25="March 1, 2021",Week26="March 15, 2021",
+              Week27="March 29, 2021")
+
+
+############################################################################
+#Data Cleaning - read needed file, and filter out needed
+
+data_clean <- read.csv("anx_Data_estimates.csv")[,-1] %>%
+  dplyr::filter(!State %in% c("DC")) %>%
+  dplyr::filter(Age %in% c("18 - 29")) %>%
+  dplyr::select(State,date_end,Response)
+
+#Spread data to make it matrix form
+data_clean_prep <- data_clean %>%
+  tidyr::spread(State,Response) %>%
+  arrange(date_end)
+
+
+# Finding which t represents sundays (what points to interpolate)
+
+lub <- as.Date(data_clean_prep$date_end)
+full_t <- seq(lub[1],lub[27],by="1 days")
+is_Sunday <-full_t[which(wday(full_t)==1)]
+
+#################################################################
+### Extrapolation
+
+my_list <- list()
+for(i in 1:length(is_Sunday)){
+  #Get date for which to interpolate
+  sunday <- is_Sunday[i]
+  
[…]

### 2021-04-11 submissiondraftNR.Rmd
trends data for "lockdown", draft of report cont.

@@
+---
 title: 'Draft of Report for Datathon'
-author: "Nathan Ryder Mantautas Rimkus, Paul Chong, Ian Taylor"
+author: "Mantautas Rimkus, Paul Chong, Nathan Ryder, Ian Taylor"
 date: '`r format(Sys.time(), "%d %B %Y")`'
@@
 
+```{r setup, include=F}
+library(knitr)
+library(tidyverse)
+library(ggplot2)
+library(gridExtra)
+library(lubridate)
+library(maps)
+
+## hide/show code
+show <- FALSE
+cac <- FALSE
+cvhei <- 7.5
+monhei <- 5
+cacpath <- "/home/nthnrydr/research_local/rmd_caches/dat_analysis_2011/"
+
+options(knitr.kable.NA = '')
+knitr::opts_chunk$set(cache.path = cacpath,
+                      cache = cac,
+                      fig.height=3.5, fig.width=5, fig.pos = "Ht",
+                      fig.path = "/home/nthnrydr/research_local/figures/",
+                      echo = show)
+                      ## base.dir = "/home/nthnrydr/research_local/rmd/dat_analysis_2011_base/",
+
+```
+
 # Introduction
@@
 Universities transitioned to partially or fully online curriculum, adding distance between student and staff communities.
-Unsurprisingly, measures for anxiety and depression have increased under the duration of the Covid-19 pandemic.
-This report examines the survey-reported changes in anxiety or depression prevalence for college-aged persons in the United States.
-Data on Covid-19 infection rates are linked with the mental health surveys.
+Unsurprisingly, measures for anxiety have increased under the duration of the Covid-19 pandemic.
+This report examines the survey-reported changes in anxiety for college-aged persons in the United States.
+Data on Covid-19 infection rates are linked with the mental health surveys via a fixed effects model.
+Google search trends are also tested for correlations with weekly anxiety measures.
 
@@
 
+The US Household Pulse Survey began in April 2020 as an attempt to monitor potential effects of the Covid pandemic on the US population.
+The survey asked questions on education, employment, food security, health, and other household characteristics that may be susceptible to change during a health emergency.
+There have been three phases to the survey, spanning from April 2020 to March 2021 with semi-weekly data available.
+For this study, Table 2a: "Symptoms of Anxiety Experienced in the Last 7 Days, by Select Characteristics" was gathered for each week of the survey.
+This table tracked the frequency that anxiety was experienced for the survey week, subjects responded with either "Not at all", "Several days", "More than half the days", or "Nearly every day".
+Additional variables, such as age, sex, race, education, etc.
+were recorded, and the survey results were split into the 50 states as well as overall US .
+
+The anxiety survey data posed a few difficulties.
+First, the survey "weeks" were at irregular intervals and an inconsistent number of days.
+For example, the first survey "week" occurred between April 23 and May 5.
+Second, the survey method used the Census Bureau's Master Address File to send emails and texts to very large samples.
+This method had high non-response rates and is likely subject to considerable bias, although this is a common feature of any survey data.
+Third, the age ranges recorded in this survey begin with "18-29".
+To demonstrate that this age range is highly relevant to university administration, data was obtained from the National Center for Education Statistics.
+Figure \@ref(fig:univpop) shows that ~30\% of 18-29 year-olds in the United States are in higher education and that of those in higher education in the US, ~75\% are between the ages 18 and 29.
+Thus the mental health characteristics of the age bracket 18-29 are will apply to the vast majority of university students.
+
+```{r, univpop, fig.cap = "The proportion of US higher education students who are between ages 18-29 (blue) and the proportion of US 18-29 year olds who are in higher education (red)."}
+ college_pop <- read.csv("~/Dropbox/Datathon/StanfordDatathonCSU/college_population_data.csv")
+ total_pop <- read.csv("~/Dropbox/Datathon/StanfordDatathonCSU//18-29_population_data.csv") %>%
+     mutate(Year = Year..August.) %>%
+     filter(Year < 2019)
+
+ ggplot(college_pop, aes(x = Year)) +
+     geom_line(aes(y = college_pop$Porportion.Inside, color = "blue")) +
+     geom_point(aes(y = college_pop$Porportion.Inside, color = "blue"))+
+     geom_line(data = total_pop, aes(y = total_pop$Proportion, color = "red")) +
+     geom_point(data = total_pop, aes(y = total_pop$Proportion, color = "red"))+
+     scale_y_continuous(breaks = c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1.00), 
+                        limits = c(0,1)) +
+     labs(x = "Year", y = "Proportion", colour = "Age Groups") +
+     theme_bw() +
+     theme(legend.position = "none")
+ ```
+
 Figures:
+
+weeks
 

### 2021-04-11 Population_Study.Rmd

---
title: "Population Study"
author: "Paul Chong"
date: "4/10/2021"
output: pdf_document
---

\colorlet{shadecolor}{gray!10}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
\newcommand{\hstart}{ \colorlet{shadecolor}{orange!20}
\begin{shaded} }
\newcommand{\hstop}{  \end{shaded} \colorlet{shadecolor}{gray!10}}

```{r}
library(ggplot2)
library(tidyr)
library(dplyr)
```

```{r}
college_pop <- read.csv("~/Datathon 2021/college_population_data.csv")
total_pop <- read.csv("~/Datathon 2021/18-29_population_data.csv")
college_pop
```

```{r, fig.cap = "The first line is the proportion of students who are between ages 18-29 (blue). The second line is the proportion of 18-29 year olds who are attending a university (red)."}

p <- ggplot(college_pop, aes(x = Year)) +
  geom_line(aes(y = college_pop$Porportion.Inside, color = "blue")) +
  geom_point(aes(y = college_pop$Porportion.Inside, color = "blue"))+
  geom_line(data = total_pop, aes(y = total_pop$Proportion, color = "red")) +
  geom_point(data = total_pop, aes(y = total_pop$Proportion, color = "red"))+
  scale_y_continuous(breaks = c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1.00), 
                     limits = c(0,1)) +
  labs(x = "Year", y = "Proportions",  
       title = "Different Measurements for University Students Aged 18-29",
       colour = "Age Groups") + 
  scale_color_manual(labels = c("Students in 18-29", "18-29 in Students"), values = c("blue", "red")) +
  theme_bw()

p
```

\hstart

University students have been found to experience mental health disorders at a disproportionate rate.  One meta-analysis found the average rate of depression in university students to be 30.6\%, when the general US population is estimated to have a 9\% average rate (REF).  The age range of most university students, 18-22, is host to large social and mental developments and is one of the more common time periods for mental disorders like anxiety to manifest (REF).  Adding the pressures of school and high-impact life choices, it is understandable that university students experience more than their fair share of mental health issues.

The year 2020 and the disease Covid-19 added further difficulty to the lives of those who are college aged.  Students who returned or were admitted to universities in fall 2020 were forced to learn new styles of learning and make sacrifices in their social lives.   Universities transitioned to partially or fully online curriculum, adding distance between student and staff communities.  Unsurprisingly, measures for anxiety and depression have increased under the duration of the Covid-19 pandemic.  This report examines the survey-reported changes in anxiety or depression prevalence for college-aged persons in the United States.  Data on Covid-19 infection rates are linked with the mental health surveys.

We decided to focus on individuals between 18-29 because this is the age group that primarily makes up the university population. Although only 30% of individuals aged between 18-29 attend universities, more than 70% of students attending universities in the United States are between 18-29. This has been a consistent trend for the past 10 years, so we can assume that the figures for 2020 is going to be approximately the same. Our team believes that because of the population proportions that we have found for university students, the following findings will apply to most of the students enrolled in universities.

\hstop

\hstart

EXPLANATION ON DATA

The first dataset we used to find university populations was a data frame we found from the National Center for Education Statistics. The dataset records total fall enrollment in degree-granting institutions by race, gender, and age of student. We extracted data from this table to create our own dataframe that contains different age populations for enrollment for every year. Each row (observation) contains the year of the semester, age groups for 14-17, 18-19, 20-21, 22-24, 25-29, 30-34, and 35 and older. Each row also contains the total enrollment for that semester, the proportion that are between ages 18-29, and the proportion that is not in that age range.

The next dataset we utilized for our study was from the United States Census Bureau Household Pulse Survey. This experimental survey was conducted to quickly and efficiently collect data on how people's lives were impacted by COVID-19 pandemic. The survey was conducted over a period of 26 weeks, where in each week there were detailed tables provided of various topics, such as education, employment, health, etc. It should be noted that the term "weeks" is not defined as a 7-day period, but rather a set period of time determined by the researchers. For example, the table for week 1 contains data for April 23 to May 5.

We focused on a table that tracked symptoms of anxiety by different characteristics, such as age, gender, marital status, education levels, etc. Our team extracted the anxiety tables for the age group 18-29 from all 26 tables for every state for each week. We then created a new data frame that contains the anxiety levels for those aged 18-29 for 26 weeks. Each individual was asked how frequent they were feeling nervous, anxious, or on the edge. They were also asked frequency of not being able to stop or control worrying. In each of these questions, the options were "Not at all", "Several days", "More than half the days", "Nearly everyday", or "Did not report".

\hstop

### 2021-04-11 submissiondraftIMT.Rmd
Included more about data, model, results.

@@
 The `plm` package in `R` allows us to fit both of these models and perform a hypothesis test to determine which fits our data better.
+
+Because COVID cases can vary largely by state due to the size of the state, we were careful in how we constructed the $\mbox{COVID}_{s,t}$ covariate.
+We used the log ratio of new cases in sequential weeks as the COVID covariate:
+$$\mbox{COVID}_{s,t} = \log\left(\frac{\text{New cases in state $s$ and week $t$}}{\text{New cases in state $s$ and week $t-1$}}\right).$$
+
+This value is appealing for three reasons.
+First, it places all states on the same scale by only considering the ratio of subsequent weeks.
+Second, by using a ratio of new covid cases it handles the exponential growth nature of the pandemic.
+And third, it works well with an autoregressive model.
+A positive value means the pandemic is currently accelerating in a state, possibly resulting in increasing anxiety relative to week $t-1$.
+Similarly, a negative value means the pandemic is slowing down, possibly resulting in decreasing anxiety relative to week $t-1$.
+
+### Results
+
+```{r getmodeldata, include=FALSE}
+library(plm)
+
+# Read in anxiety data, covid data (whole US) and covid data (by state)
+anxiety <- read.csv("anx_rate_ext.csv")
+covid.us <- read.csv("covid-data/weekly-us.csv")
+covid <- read.csv("covid-data/weekly-states.csv")
+
+# Create logitudinal long form data frame
+paneldata <- data.frame(
+  State = covid$State,
+  Date = covid$EndofWeek,
+  CasesPerDay = covid$Cases,
+  LogCasesRatio = covid$LogRatio
+)
+for (i in 1:nrow(paneldata)) {
+  s <- paneldata[i, "State"]
+  d <- paneldata[i, "Date"]
+
+  if (d %in% anxiety$date_end) {
+    paneldata[i,"anx"] <- anxiety[anxiety$date_end == d, s]
+  }
+}
+
+paneldata <- paneldata[paneldata$Date %in% anxiety$date_end,] # Only dates with an anxiety measurement
+```
+
+First, we have the output from the fixed effects model:
+```{r runfixedmodel}
+# Model: panel model by state with anxiety as response regressed on covid
+fit.states.fixed <- plm(anx ~ LogCasesRatio + lag(anx), paneldata, 
+                        index=c("State", "Date"), model="within")
+summary(fit.states.fixed)
+```
+
+We can see that both the lagged anxiety variable and LogCasesRatio are significant predictors.
+The value of the coefficient with $\mbox{COVID}_{s,t}$ shows that if new cases were to double from one week to the next, we expect our response anxiety odds to increase by `r round(fit.states.fixed$coefficients[1] * log(2), digits=4)`.
+
+The `plm` package also allows us to run a Hausman hypothesis test to see whether a fixed effects or random effects model is a better fit for our data.
+
+```{r runrandommodel}
+fit.states.random <- plm(anx ~ LogCasesRatio + lag(anx), paneldata, 
+                         index=c("State", "Date"), model="random")
+phtest(fit.states.fixed, fit.states.random)
+```
+
+The small p-value indicates that the null hypothesis of a random effects model should be rejected, and so our results from the fixed effects model are a better fit.

### 2021-04-11 Submission template_Mantautas.Rmd
ONE MORE TIME

@@
 
-```{r}
-require(rmgarch)
+Description of Permutation Test.
 
-````
+As it was seen in the given map, there are some spatial-temporal trends between the difference in anxiety ratio and differences in Google Search trends for Covid.
+For the sake of this project, the joint significance of correlations of Colorado (where the team is based) and of California was tested using a permutation test.
+The statistics $T$ were calculated as
+
+$$
+T=corr^2_{\text{California}}+corr^2_{\text{Colorado}}
+$$
+
+The sampling distribution was derived using a permutation sampling, wherein each iteration of each states differences of anxiety ratio is assigned to randomized states differences in google trends and sampling distribution was obtained as $T_{1},\ldots, T_{1000}$.
+
+We obtained $p=0.12$, thus the hypothesis of no significance cannot be rejected.
+Thus notice that taking into account the nature of data (noisy, interpolation), etc.
+One can argue that $p-value$ around significance $0.1$ that there is potential for more research in this area.

### 2021-04-11 anxiety-covid-ts-model.R
Fixed typos

@@ -1,7 +1,6 @@
 # This file runs a time series model for anxiety of 18-29 year olds with
 # covid cases as a predictor
 
-dplyr
 library(plm)
 
 # Read in anxiety data, covid data (whole US) and covid data (by state)
@@ -21,7 +20,7 @@ search <- search %>%
   )
 
 # Join together data sources into data frame for linear model
-dates <- anx$date_end
+dates <- anxiety$date_end
 
 modeldata <- data.frame(
   Week = dates,

### 2021-04-11 Supplemental R code/Permutation.R
putting R codes into one place

@@ -1,0 +1,61 @@
+#############################################################################
+####Running permutation test for correlation
+####Output:
+###############
+
+#############################################################################
+#Libraries 
+
+library(tidyverse)
+library(plyr)
+library(readr)
+library("lubridate")
+library(fda)
+
+#############################################################################
+
+data_needed <- read.csv("diffsdf.csv") %>% na.omit()
+ 
+# Calculate statistics value for California and Colorado
+
+Stats <- data_needed %>%
+  dplyr::group_by(State) %>%
+  dplyr::filter(State==c("CA","CO")) %>% 
+  dplyr::summarise(corr=cor(hitdiffs,anxdiffs)) %>%
+  dplyr::summarise(mean(corr*corr)) %>%
+  as.numeric()
+
+# Prepare data frame to use for joining
+anx_not_mix <- data_needed %>%
+  dplyr::select(State,date_end,anxdiffs)
+
+#Prepare data frame that gonna be mixed
+to_mix_hitdiffs <- data_needed %>%
+  dplyr::select(State,hitdiffs,date_end) %>%
+  tidyr::spread(date_end,hitdiffs)
+
+#Save statitics in here
+N <- 1000
+permut <- 1:nrow(to_mix_hitdiffs)
+save_stats_permut <- rep(0,N)
+
+#######################################################33
+#Run permutation test
+for(i in 1:N){
+  mixing <- to_mix_hitdiffs
+  mixing$State <- to_mix_hitdiffs$State[sample(permut,replace=FALSE)]
+  save_stats_permut[i] <- anx_not_mix %>%
+  left_join(mixing %>%
+  gather("date_end","hitdiffs",-State),by = c("State", "date_end")) %>% 
+    na.omit() %>%
+    dplyr::filter(State==c("CA","CO")) %>% 
+    dplyr::group_by(State) %>%
+    dplyr::summarise(corr=(cor(hitdiffs,anxdiffs))) %>%
+    dplyr::summarise(mean(corr*corr)) %>%
+    as.numeric()
+}
+
+####################################
+pvalue=mean(save_stats_permut > Stats)
+
+

### 2021-04-11 Supplemental R code/anxiety-covid-ts-model.R

@@ -1,0 +1,134 @@
+# This file runs a time series model for anxiety of 18-29 year olds with
+# covid cases as a predictor
+
+library(plm)
+
+# Read in anxiety data, covid data (whole US) and covid data (by state)
+anxiety <- read.csv("anx_rate_ext.csv")
+covid.us <- read.csv("covid-data/weekly-us.csv")
+covid <- read.csv("covid-data/weekly-states.csv")
+search <- read.csv("statetrends.csv")
+
+########## Add diffs to search data
+`%>%` <- dplyr::`%>%`
+search <- search %>%
+  dplyr::arrange(location, date_end) %>%
+  dplyr::group_by(location) %>%
+  dplyr::mutate(
+    diff = hits - dplyr::lag(hits),
+    logratio = log(hits / dplyr::lag(hits))
+  )
+
+# Join together data sources into data frame for linear model
+dates <- anxiety$date_end
+
+modeldata <- data.frame(
+  Week = dates,
+  anx = anxiety$US
+)
+for (i in 1:nrow(modeldata)) {
+  d <- modeldata[i, "Week"]
+  d.lag1 <- as.character(as.Date(d) - 7)
+  d.lag2 <- as.character(as.Date(d) - 14)
+  
+  # Anxiety from the previous week(s)
+  if (d.lag1 %in% anxiety$date_end) { 
+    modeldata[i, "anx.lag1"] <- anxiety[anxiety$date_end == d.lag1, "US"]
+  }
+  if (d.lag2 %in% anxiety$date_end) { 
+    modeldata[i, "anx.lag2"] <- anxiety[anxiety$date_end == d.lag2, "US"]
+  }
+  
+  # Covid cases (log ratio change) in current week and previous week(s)
+  if (d %in% covid.us$EndofWeek) {
+    modeldata[i, "covid"] <- covid.us[covid.us$EndofWeek == d, "LogRatio"]
+  }
+  if (d.lag1 %in% covid.us$EndofWeek) {
+    modeldata[i, "covid.lag1"] <- covid.us[covid.us$EndofWeek == d.lag1, "LogRatio"]
+  }
+  if (d.lag1 %in% covid.us$EndofWeek) {
+    modeldata[i, "covid.lag2"] <- covid.us[covid.us$EndofWeek == d.lag2, "LogRatio"]
+  }
+}
+modeldata$Election <- modeldata$Week %in% c("2020-11-01", "2020-11-08")
+
+
+################################################################################
+# Run linear models
+################################################################################
+
[…]

### 2021-04-11 Supplemental R code/plot-and-save-covid-data.R

@@ -1,0 +1,67 @@
+library(lubridate)
+library(dplyr)
+library(ggplot2)
+library(zoo)
+
+# Read covid data
+covid <- read.csv("covid-data/United_States_COVID-19_Cases_and_Deaths_by_State_over_Time.csv")
+
+states <- c("AK", "AL", "AR", "AZ", "CA", "CO", "CT", "DE", "FL", "GA",
+            "HI", "IA", "ID", "IL", "IN", "KS", "KY", "LA", "MA", "MD",
+            "ME", "MI", "MN", "MO", "MS", "MT", "NC", "ND", "NE", "NH",
+            "NJ", "NM", "NV", "NY", "OH", "OK", "OR", "PA", "RI", "SC",
+            "SD", "TN", "TX", "UT", "VA", "VT", "WA", "WI", "WV", "WY",
+            "NYC") # Include NYC, to be combined with NY
+
+# Process dates for sorting
+covid <- data.frame(
+  Date = as.Date(strptime(covid$submission_date, "%m/%d/%Y")),
+  State = covid$state,
+  Cases = covid$new_case
+)
+covid <- covid[covid$State %in% states,]
+# Combine NYC with NY by taking first 2 substring
+covid$State <- substr(covid$State, 1, 2)
+covid <- aggregate(Cases ~ Date + State, covid, sum)
+# Add week
+covid$EndofWeek <- covid$Date - (wday(covid$Date) - 1) + 7 * (wday(covid$Date) != 1)
+
+
+# Aggregate to weekly and calculate sequential log ratios
+covid.weekly <- covid %>%
+  group_by(State, EndofWeek) %>%
+  summarize(Cases = mean(Cases))
+covid.weekly <- covid.weekly %>%
+  arrange(State, EndofWeek) %>%
+  group_by(State) %>%
+  mutate(LogRatio = log(Cases / lag(Cases)))
+covid.weekly$LogRatio <- replace(covid.weekly$LogRatio, is.infinite(covid.weekly$LogRatio), NaN)
+  
+# Plot each state as a series
+ggplot(covid.weekly, aes(x=EndofWeek, y=LogRatio)) + geom_line(aes(color=State))
+
+#############################################################################
+### Aggregate to the US
+
+# Aggregate
+covid.us <- aggregate(Cases ~ Date + EndofWeek, covid, sum)
+
+# Aggregate to weekly and caluculate sequential log ratios
+covid.us.weekly <- covid.us %>%
[…]

### 2021-04-11 Anxiety_Data_Interpolation.R
Delete Anxiety_Data_Interpolation.R

@@ -1,89 +1,0 @@
-#############################################################################
-####This code for anxiety estimates interpolation
-####Output: File anx_rate_ext.csv
-###############
-
-############################################################################
-#Libraries 
-
-library(tidyverse)
-library(plyr)
-library(readr)
-library("lubridate")
-library(fda)
-
-###########################################################################
-#Setting up weeks dates. 
-
-dates_end<- c(Week1="May 5, 2020",Week2="May 12, 2020",Week3="May 19, 2020",
-              Week4="May 26, 2020",Week5="June 2, 2020",Week6="June 9, 2020",
-              Week7="June 16, 2020",Week8="June 23, 2020",Week9="June 30, 2020",
-              Week10="July 7, 2020",Week11="July 14, 2020",Week12="July 21, 2020",
-              Week13="August 31, 2020",Week14="September 14, 2020",
-              Week15="September 28, 2020",Week16="October 12, 2020",
-              Week17="October 26, 2020",Week18="November 9, 2020",
-              Week19="November 23, 2020",Week20="December 7, 2020",
-              Week21="December 21, 2020", Week22="January 18, 2021",
-              Week23="February 1, 2021",Week24="February 15, 2021",
-              Week25="March 1, 2021",Week26="March 15, 2021",
-              Week27="March 29, 2021")
-
-
-############################################################################
-#Data Cleaning - read needed file, and filter out needed
-
-data_clean <- read.csv("anx_Data_estimates.csv")[,-1] %>%
-  dplyr::filter(!State %in% c("DC")) %>%
-  dplyr::filter(Age %in% c("18 - 29")) %>%
-  dplyr::select(State,date_end,Response)
-
-#Spread data to make it matrix form
-data_clean_prep <- data_clean %>%
-  tidyr::spread(State,Response) %>%
-  arrange(date_end)
-
-
-# Finding which t represents sundays (what points to interpolate)
-
-lub <- as.Date(data_clean_prep$date_end)
-full_t <- seq(lub[1],lub[27],by="1 days")
-is_Sunday <-full_t[which(wday(full_t)==1)]
-
-#################################################################
-### Extrapolation
-
-my_list <- list()
-for(i in 1:length(is_Sunday)){
-  #Get date for which to interpolate
-  sunday <- is_Sunday[i]
-  
[…]

### 2021-04-11 Anxiety_Data_Read_And_Prepare.R
Delete Anxiety_Data_Read_And_Prepare.R

@@ -1,260 +1,0 @@
-#############################################################################
-####This code takes data from Household survey and prepares it for modeling
-####Output: Files anx_Data_estimates.csv" and anx_Data_error.csv
-###############
-
-####Libraries
-library(readxl)
-library(plyr)
-library(tidyverse)
-
-##############################################################################
-####Special functions
-
-  #Function for reading all excel sheets
-read_excel_allsheets <- function(filename, tibble = FALSE) {
-  # I prefer straight data.frames
-  # but if you like tidyverse tibbles (the default with read_excel)
-  # then just pass tibble = TRUE
-  sheets <- readxl::excel_sheets(filename)
-  x <- lapply(sheets, function(X) readxl::read_excel(filename, sheet = X))
-  if(!tibble) x <- lapply(x, as.data.frame)
-  names(x) <- sheets
-  x
-}
-
-  #Settings dates for each each week 
-dates_beg<- c(Week1="April 23, 2020",
-              Week2="May 7, 2020",
-              Week3="May 14, 2020",
-              Week4="May 21, 2020",
-              Week5="May 28, 2020",
-              Week6="June 4, 2020",
-              Week7="June 11, 2020",
-              Week8="June 18, 2020",
-              Week9="June 25, 2020",
-              Week10="July 2, 2020",
-              Week11="July 9, 2020",
-              Week12="July 16, 2020",
-              Week13="August 19, 2020",
-              Week14="September 2, 2020",
-              Week15="September 16, 2020",
-              Week16="September 30, 2020",
-              Week17="October 14, 2020",
-              Week18="November 9, 2020",
-              Week19="November 23, 2020",
-              Week20="December 7, 2020",
-              Week21="December 21, 2020",
-              Week22="January 18, 2021",
-              Week23="February 1, 2021",
-              Week24="February 15, 2021",
-              Week25="March 1, 2021",
-              Week26="March 15, 2021",
-              Week27="March 29, 2021")
-
-dates_end<- c(Week1="May 5, 2020",
-              Week2="May 12, 2020",
-              Week3="May 19, 2020",
-              Week4="May 26, 2020",
-              Week5="June 2, 2020",
-              Week6="June 9, 2020",
-              Week7="June 16, 2020",
-              Week8="June 23, 2020",
[…]

### 2021-04-11 anxiety-covid-ts-model.R
Delete anxiety-covid-ts-model.R

@@ -1,134 +1,0 @@
-# This file runs a time series model for anxiety of 18-29 year olds with
-# covid cases as a predictor
-
-library(plm)
-
-# Read in anxiety data, covid data (whole US) and covid data (by state)
-anxiety <- read.csv("anx_rate_ext.csv")
-covid.us <- read.csv("covid-data/weekly-us.csv")
-covid <- read.csv("covid-data/weekly-states.csv")
-search <- read.csv("statetrends.csv")
-
-########## Add diffs to search data
-`%>%` <- dplyr::`%>%`
-search <- search %>%
-  dplyr::arrange(location, date_end) %>%
-  dplyr::group_by(location) %>%
-  dplyr::mutate(
-    diff = hits - dplyr::lag(hits),
-    logratio = log(hits / dplyr::lag(hits))
-  )
-
-# Join together data sources into data frame for linear model
-dates <- anxiety$date_end
-
-modeldata <- data.frame(
-  Week = dates,
-  anx = anxiety$US
-)
-for (i in 1:nrow(modeldata)) {
-  d <- modeldata[i, "Week"]
-  d.lag1 <- as.character(as.Date(d) - 7)
-  d.lag2 <- as.character(as.Date(d) - 14)
-  
-  # Anxiety from the previous week(s)
-  if (d.lag1 %in% anxiety$date_end) { 
-    modeldata[i, "anx.lag1"] <- anxiety[anxiety$date_end == d.lag1, "US"]
-  }
-  if (d.lag2 %in% anxiety$date_end) { 
-    modeldata[i, "anx.lag2"] <- anxiety[anxiety$date_end == d.lag2, "US"]
-  }
-  
-  # Covid cases (log ratio change) in current week and previous week(s)
-  if (d %in% covid.us$EndofWeek) {
-    modeldata[i, "covid"] <- covid.us[covid.us$EndofWeek == d, "LogRatio"]
-  }
-  if (d.lag1 %in% covid.us$EndofWeek) {
-    modeldata[i, "covid.lag1"] <- covid.us[covid.us$EndofWeek == d.lag1, "LogRatio"]
-  }
-  if (d.lag1 %in% covid.us$EndofWeek) {
-    modeldata[i, "covid.lag2"] <- covid.us[covid.us$EndofWeek == d.lag2, "LogRatio"]
-  }
-}
-modeldata$Election <- modeldata$Week %in% c("2020-11-01", "2020-11-08")
-
-
-################################################################################
-# Run linear models
-################################################################################
-
[…]

### 2021-04-11 plot-and-save-covid-data.R
Delete plot-and-save-covid-data.R

@@ -1,67 +1,0 @@
-library(lubridate)
-library(dplyr)
-library(ggplot2)
-library(zoo)
-
-# Read covid data
-covid <- read.csv("covid-data/United_States_COVID-19_Cases_and_Deaths_by_State_over_Time.csv")
-
-states <- c("AK", "AL", "AR", "AZ", "CA", "CO", "CT", "DE", "FL", "GA",
-            "HI", "IA", "ID", "IL", "IN", "KS", "KY", "LA", "MA", "MD",
-            "ME", "MI", "MN", "MO", "MS", "MT", "NC", "ND", "NE", "NH",
-            "NJ", "NM", "NV", "NY", "OH", "OK", "OR", "PA", "RI", "SC",
-            "SD", "TN", "TX", "UT", "VA", "VT", "WA", "WI", "WV", "WY",
-            "NYC") # Include NYC, to be combined with NY
-
-# Process dates for sorting
-covid <- data.frame(
-  Date = as.Date(strptime(covid$submission_date, "%m/%d/%Y")),
-  State = covid$state,
-  Cases = covid$new_case
-)
-covid <- covid[covid$State %in% states,]
-# Combine NYC with NY by taking first 2 substring
-covid$State <- substr(covid$State, 1, 2)
-covid <- aggregate(Cases ~ Date + State, covid, sum)
-# Add week
-covid$EndofWeek <- covid$Date - (wday(covid$Date) - 1) + 7 * (wday(covid$Date) != 1)
-
-
-# Aggregate to weekly and calculate sequential log ratios
-covid.weekly <- covid %>%
-  group_by(State, EndofWeek) %>%
-  summarize(Cases = mean(Cases))
-covid.weekly <- covid.weekly %>%
-  arrange(State, EndofWeek) %>%
-  group_by(State) %>%
-  mutate(LogRatio = log(Cases / lag(Cases)))
-covid.weekly$LogRatio <- replace(covid.weekly$LogRatio, is.infinite(covid.weekly$LogRatio), NaN)
-  
-# Plot each state as a series
-ggplot(covid.weekly, aes(x=EndofWeek, y=LogRatio)) + geom_line(aes(color=State))
-
-#############################################################################
-### Aggregate to the US
-
-# Aggregate
-covid.us <- aggregate(Cases ~ Date + EndofWeek, covid, sum)
-
-# Aggregate to weekly and caluculate sequential log ratios
-covid.us.weekly <- covid.us %>%
[…]

### 2021-04-11 submissiondraftNR.Rmd
next draft

@@
 ---
-title: 'Draft of Report for Datathon'
+title: Title
 author: "Mantautas Rimkus, Paul Chong, Nathan Ryder, Ian Taylor"
@@
 library(ggplot2)
+library(gghighlight)
 library(gridExtra)
+library(plm)
+library(gtrendsR)
 library(lubridate)
@@
 
+## file location prefix
+datadir <- "/home/nthnrydr/Dropbox/Datathon/StanfordDatathonCSU/"
 ```
+Team No: 202
+
+Track No.: 2.
+
+Link to GitHub Repository: https://github.com/MantautasRimkus/StanfordDatathonCSU
 
@@
 University students have been found to experience mental health disorders at a disproportionate rate.
-One meta-analysis found the average rate of depression in university students to be 30.6\%, when the general US population is estimated to have a 9\% average rate (REF).
-The age range of most university students, 18-22, is host to large social and mental developments and is one of the more common time periods for mental disorders like anxiety to manifest (REF).
+One meta-analysis found the average rate of depression in university students to be 30.6\%, when the general US population is estimated to have a 9\% average rate \textbf{REF}.
+The age range of most university students, 18-22, is host to large social and mental developments and is one of the more common time periods for mental disorders like anxiety to manifest \textbf{REF}.
 Adding the pressures of school and high-impact life choices, it is understandable that university students experience more than their fair share of mental health issues.
@@
 
-The US Household Pulse Survey began in April 2020 as an attempt to monitor potential effects of the Covid pandemic on the US population.
+The \textbf{REF} US Household Pulse Survey began in April 2020 as an attempt to monitor potential effects of the Covid pandemic on the US population.
 The survey asked questions on education, employment, food security, health, and other household characteristics that may be susceptible to change during a health emergency.
@@
 First, the survey "weeks" were at irregular intervals and an inconsistent number of days.
-For example, the first survey "week" occurred between April 23 and May 5.
+The irregularity of the intervals can be seen by some gaps in observations in Figure \@ref{fig:usanx}.
 Second, the survey method used the Census Bureau's Master Address File to send emails and texts to very large samples.
@@
 Third, the age ranges recorded in this survey begin with "18-29".
-To demonstrate that this age range is highly relevant to university administration, data was obtained from the National Center for Education Statistics.
+To demonstrate that this age range is highly relevant to university administration, data was obtained from the National Center for Education Statistics \textbf{REF}.
 Figure \@ref(fig:univpop) shows that ~30\% of 18-29 year-olds in the United States are in higher education and that of those in higher education in the US, ~75\% are between the ages 18 and 29.
@@
 ```{r, univpop, fig.cap = "The proportion of US higher education students who are between ages 18-29 (blue) and the proportion of US 18-29 year olds who are in higher education (red)."}
- college_pop <- read.csv("~/Dropbox/Datathon/StanfordDatathonCSU/college_population_data.csv")
- total_pop <- read.csv("~/Dropbox/Datathon/StanfordDatathonCSU//18-29_population_data.csv") %>%
+ college_pop <- read.csv(paste0(datadir, "college_population_data.csv"))
+ total_pop <- read.csv(paste0(datadir, "18-29_population_data.csv")) %>%
      mutate(Year = Year..August.) %>%
@@
 
+To make a single value metric for anxiety levels, the survey estimates for the last two responses in the survey ("More than half the days" and "Nearly every day") were summed and divided by the sum of the first two responses("Not at all" and "Several days").
+The resulting metric is the ratio of survey respondents who felt anxious more than half the time divided by respondents who felt anxious less than half the time.
+Furthermore, the anxiety survey data came with standard error estimates for each response.
+Using error propagation formulas as in \textbf{REF}, upper-bound standard error estimates could be produced.
+Figure \@ref(fig:usanx) shows the anxiety ratio metric for 18-29 year-olds  plotted over the survey period.
+
+```{r, usanx, fig.cap="Anxiety ratio (more than half the time)/(less than half the time) of 18-29 year-olds plotted over survey period with standard error estimates at each observation."}
+## read in surveyed anxiety estimates and errors
+anx_est <- read.csv(paste0(datadir, "anx_Data_estimates.csv"))
+anx_err <- read.csv(paste0(datadir, "anx_Data_error.csv"))
+
+## make correlations to be used in std error estimates
+anx_est <- anx_est %>%
+    group_by(State) %>%
+    mutate(corMoreNearly = cor(More.than.half.the.days, Nearly.every.day),
+           corNotSeveral = cor(Not.at.all, Several.days)) %>%
+    ungroup()
+
+## make std error estimate (should be overestimate since this doesn't include covariance terms that would be subtracted)
+anx_err <- anx_err %>%
+    mutate(Response =  anx_est$Response * sqrt(
+    (More.than.half.the.days^2 + 2*Nearly.every.day*More.than.half.the.days*anx_est$corMoreNearly + Nearly.every.day^2)/(anx_est$More.than.half.the.days + anx_est$Nearly.every.day)^2 + 
+    (Several.days^2 + 2*Several.days*Not.at.all*anx_est$corNotSeveral + Not.at.all^2)/(anx_est$Several.days + anx_est$Not.at.all)^2)) %>%
+    mutate(Upper = anx_est$Response + Response, 
+           Lower = anx_est$Response - Response)
+
+## plot us overall with std error bars
+age <- "18 - 29"
+state <- "US"
+ggplot(filter(anx_est, State == state & Age == age),
+       aes(x = as.Date(date_end), y = Response)) +
+    geom_line(alpha = .8) +
+    geom_errorbar(aes(ymax = Upper, ymin = Lower), alpha = .7,
+                  data = filter(anx_err, State == state & Age == age)) +
+    theme_bw() +
+    xlab("") +
+    ylab("Anxiety Ratio") +
+    scale_y_continuous(limits = c(.4, 1.2))
+```
+
+The Household Pulse Survey occurred in three phases, which can be seen in Figure \@ref(fig:usanx) as gaps in observed time points.
+The anxiety ratio for 18-29 year-olds had an overall upwards trend for the year period, and the jumps in summer and winter may be related to spikes in Covid infections.
+In November 2020 a large increase in anxiety can be easily related to the 2020 presidential election.
+
+As described in the technical documentation (LINK), the weeks of the Household Pulse Survey do not represent strict intervals of 7 days.
+For example, "Week 1" refers to data that was collected between April 23rd and May 5th, 2020, (13 days), but Week 9 refers to data that was collected between June 25th and June 30th, 2020 (6 days).
+The time points $s_i$ will be defined as the end date of each Week collection.
+
+Notice that the $s_i$ are sparse and lead to unequal spaces between data points.
+To ensure univariate differences between time points, and be able to implement the models described later, interpolation was used with new time points $t_1,\ldots,t_{47}$ representing Sundays between May 05, 2020, and March 29, 2021.
+To calculate each $X_{t_i}$, where $s_{j-1}<t_i\leq s_{j}$ the formula :
+
+$$
+X_{t_i}=\frac{d(s_{j-1},t_i)}{d(s_{j-1},s_{j})}X_{s_j} + \frac{d(t_i,s_{j})}{d(s_{j-1},s_{j})}X_{s_{j-1}} \text{ was used.}
+$$
+
+Figure \@ref(fig:coanx) depicts raw data $X_{s_1},\ldots,X_{s_27}$ and interpolated data $X_{t_1},\ldots, X_{t_j}$ for anxiety ratios in Colorado.
+Figure \@ref(fig:statanx) shows the interpolated anxiety ratios for all states with Colorado and California emphasized.
+
+```{r, coanx, fig.cap="Anxiety Ratio in Colorado. Black - original data points, Red - interpolated data with equal time spaces", message=FALSE }
+data_raw <- read.csv(paste0(datadir, "anx_Data_estimates.csv"))
+data_inter <- read.csv(paste0(datadir, "anx_rate_ext.csv"))
+
+data_raw %>%
+  dplyr::filter(State %in% c("CO")) %>%
+  dplyr::filter(Age %in% c("18 - 29")) %>%
+  dplyr::select(State,date_end,Response) %>%
+  ggplot(aes(x=as.Date(date_end),y=Response)) +
+  geom_line() +
+  geom_point() +
+  geom_line(data=data_inter[,c(2,8)],aes(x=as.Date(date_end),y=CO),col="red") +         geom_point(data=data_inter[,c(2,8)],aes(x=as.Date(date_end), y=CO),col="red") + 
+   theme_bw() +
+   ylab("Anxiety ratio") +
+   xlab("Date")
+```
+
+```{r, statanx, fig.cap="Interpolated anxiety ratios for all states with emphasis for Colorado and California", message=FALSE, warning=FALSE}
+data_inter[,-1] %>%
+   tidyr::gather("State","Value",-date_end) %>%
+   dplyr::filter(!State=="US") %>%
+   ggplot(aes(x=as.Date(date_end),y=Value,col=State)) +
+   geom_line() +
+   gghighlight(State %in% c("CO","CA")) +
+   theme_bw() +
+   ylab("Anxiety ratio") +
+   xlab("Date")
+```
+
+# Modeling
+
+## Predicting Anxiety from COVID Cases Using a Panel Model
+
+We are searching for a way for school administrators to anticipate increased anxiety among their students.
+In 2020 and today, anecdotally, COVID is a significant source of anxiety.
+We used a panel model to find the predictiveness of covid cases on anxiety in 18-29 year olds.
+
+Panel models are used for regression when the response and predictor variables are longitudinal.
+In this case, we treat US states as entities and have longitudinal values for both anxiety and new COVID cases.
+
+For states $s=1,\dots, 50$ and time periods $t=1,\dots,T$, we use the model
+
+$$
+\mbox{anx}_{s,t} = \alpha_s + \beta_1 \mbox{anx}_{s,t-1} + \beta_2 \mbox{COVID}_{s,t} + \varepsilon_{s,t}
+$$
+
+This model allows for state-specific intercepts ($\alpha_s$), autocorrelation in anxiety within a state ($\beta_1$), and an effect of COVID cases on anxiety ($\beta_1$).
+
+Alternatively, we can have a random effect for the state-specific intercepts,
+\begin{align*}
+\mbox{anx}_{s,t} &= \alpha + a_s + \beta_1 \mbox{anx}_{s,t-1} + \beta_2 \mbox{COVID}_{s,t} + \varepsilon_{s,t} \\
+a_s &\sim N(0, \sigma^2_a)
+\end{align*}
+
+The `plm` package in `R` allows us to fit both of these models and perform a hypothesis test to determine which fits our data better.
+
+Because COVID cases can vary largely by state due to the size of the state, we were careful in how we constructed the $\mbox{COVID}_{s,t}$ covariate.
+We used the log ratio of new cases in sequential weeks as the COVID covariate:
+$$\mbox{COVID}_{s,t} = \log\left(\frac{\text{New cases in state $s$ and week $t$}}{\text{New cases in state $s$ and week $t-1$}}\right).$$
+
+This value is appealing for three reasons.
+First, it places all states on the same scale by only considering the ratio of subsequent weeks.
+Second, by using a ratio of new covid cases it handles the exponential growth nature of the pandemic.
+And third, it works well with an autoregressive model.
+A positive value means the pandemic is currently accelerating in a state, possibly resulting in increasing anxiety relative to week $t-1$.
+Similarly, a negative value means the pandemic is slowing down, possibly resulting in decreasing anxiety relative to week $t-1$.
+
+### Results
+
+```{r getmodeldata, include=FALSE, eval=FALSE}
+library(plm)
+
+# Read in anxiety data, covid data (whole US) and covid data (by state)
+anxiety <- read.csv(paste0(datadir, "anx_rate_ext.csv"), stringsAsFactors = FALSE)
+covid.us <- read.csv(paste0(datadir, "covid-data/weekly-us.csv"), stringsAsFactors = FALSE)
+covid <- read.csv(paste0(datadir, "covid-data/weekly-states.csv"), stringsAsFactors = FALSE)
+
+# Create logitudinal long form data frame
+paneldata <- data.frame(
+  State = covid$State,
+  Date = covid$EndofWeek,
+  CasesPerDay = covid$Cases,
+  LogCasesRatio = covid$LogRatio
+)
+for (i in 1:nrow(paneldata)) {
+  s <- paneldata[i, "State"]
+  d <- paneldata[i, "Date"]
+
+  if (d %in% anxiety$date_end) {
+    paneldata[i,"anx"] <- anxiety[anxiety$date_end == d, s]
+  }
+}
+
+paneldata <- paneldata[paneldata$Date %in% anxiety$date_end,] # Only dates with an anxiety measurement
+```
+
+First, we have the output from the fixed effects model:
+```{r runfixedmodel}
+## read df
+paneldata <- read.csv(paste0(datadir, "panel-model-data.csv"))
+
+# Model: panel model by state with anxiety as response regressed on covid
+fit.states.fixed <- plm(anx ~ LogCasesRatio + lag(anx), paneldata, 
+                        index=c("State", "Date"), model="within")
+summary(fit.states.fixed)
+```
+
+We can see that both the lagged anxiety variable and LogCasesRatio are significant predictors.
+The value of the coefficient with $\mbox{COVID}_{s,t}$ shows that if new cases were to double from one week to the next, we expect our response anxiety odds to increase by `r round(fit.states.fixed$coefficients[1] * log(2), digits=4)`.
+
+The `plm` package also allows us to run a Hausman hypothesis test to see whether a fixed effects or random effects model is a better fit for our data.
+
+```{r runrandommodel}
+fit.states.random <- plm(anx ~ LogCasesRatio + lag(anx), paneldata, 
+                         index=c("State", "Date"), model="random")
+phtest(fit.states.fixed, fit.states.random)
+```
+
+The small p-value indicates that the null hypothesis of a random effects model should be rejected, and so our results from the fixed effects model are a better fit.
+
+## Associating Google Trends Data with Anxiety Surveys
+
+The search engine Google makes its search data available through Google Trends at https://trends.google.com/trends/.
+Using the package `gtrendsr`, relative search popularity over time of a specified keyword can be obtained for a variety of regions.
+Search popularity of the keywords "lockdown" and "covid" was gathered for the study period at US and state levels.
+Figure \@ref(fig:gtrendus) shows the search popularity (0-100) of the word "covid" in the US from April 2020 to March 2021.
+
+```{r, getgtrends, }
+## start and ending time points
+start <- min(as.Date(anxiety$date_end))
+end <- max(as.Date(anxiety$date_end))
+
+## get interest over time for US
+covidint <- gtrends("covid", geo = "US", time = paste(start, end, sep = " "), onlyInterest = TRUE)[[1]]
+```
+
+```{r, gtrendus, fig.cap="Google Trends search popularity of "covid" over time.}
+ggplot(data = covidint, aes(x = as.Date(date), y = hits)) +
+    xlab("") +
+    ylab("Search Popularity") +
+    geom_line(alpha = .8)
+```
+
+## plot the trends together
+ggplot(data = anx_est_interp, aes(x = as.Date(date_end), y = US*100)) +
+    geom_line(color = "blue") +
+    geom_line(data = covidint, aes(x = as.Date(date), y = hits))
+
+## plot differences
+anx_diffs <- data.frame(date = anx_est_interp$date_end[-length(anx_est_interp$date_end)], 
+                        USdiffs = diff(anx_est_interp$US))
+
+covidint_diffs <- data.frame(date = covidint$date[-length(covidint$date)], 
+                             hitdiffs = log(covidint$hits[-length(covidint$hits)]) - log(covidint$hits[-1]))
+
+ggplot(data = anx_diffs, aes(x = as.Date(date), y = USdiffs)) +
+    geom_line(color = "blue") +
+    geom_line(data = covidint_diffs, aes(x = as.Date(date), y = hitdiffs))
+
 Figures:

### 2021-04-11 submissiondraftNR.Rmd
final draft?

@@
 ---
-title: Title
-author: "Mantautas Rimkus, Paul Chong, Nathan Ryder, Ian Taylor"
-date: '`r format(Sys.time(), "%d %B %Y")`'
 output:
-  bookdown::html_document2: 
-    number_sections: true
-    toc: yes
+  bookdown::pdf_document2: 
+    number_sections: false
+    toc: no
 
@@
 
+# Leveraging Covid and Trends Data to Inform University Policy on Anxiety
+## Mantautas Rimkus, Paul Chong, Nathan Ryder, Ian Taylor
+
 Link to GitHub Repository: https://github.com/MantautasRimkus/StanfordDatathonCSU
@@
 University students have been found to experience mental health disorders at a disproportionate rate.
-One meta-analysis found the average rate of depression in university students to be 30.6\%, when the general US population is estimated to have a 9\% average rate \textbf{REF}.
-The age range of most university students, 18-22, is host to large social and mental developments and is one of the more common time periods for mental disorders like anxiety to manifest \textbf{REF}.
+One meta-analysis found the average rate of depression in university students to be 30.6\%, when the general US population is estimated to have a 9\% average rate.
+The age range of most university students, 18-22, is host to large social and mental developments and is one of the more common time periods for mental disorders like anxiety to manifest.
 Adding the pressures of school and high-impact life choices, it is understandable that university students experience more than their fair share of mental health issues.
@@
 
+University administration is concerned with student mental health.
+This study examines relationships of age-group specific anxiety surveys with data on Covid cases or Google search trends.
+Notable relationships may be useful for the monitoring the mental health of university student populations.
+
 # Data
 
-The \textbf{REF} US Household Pulse Survey began in April 2020 as an attempt to monitor potential effects of the Covid pandemic on the US population.
+The US Household Pulse Survey began in April 2020 as an attempt to monitor potential effects of the Covid pandemic on the US population.
 The survey asked questions on education, employment, food security, health, and other household characteristics that may be susceptible to change during a health emergency.
@@
 Third, the age ranges recorded in this survey begin with "18-29".
-To demonstrate that this age range is highly relevant to university administration, data was obtained from the National Center for Education Statistics \textbf{REF}.
+To demonstrate that this age range is highly relevant to university administration, data was obtained from the National Center for Education Statistics.
 Figure \@ref(fig:univpop) shows that ~30\% of 18-29 year-olds in the United States are in higher education and that of those in higher education in the US, ~75\% are between the ages 18 and 29.
@@
 Furthermore, the anxiety survey data came with standard error estimates for each response.
-Using error propagation formulas as in \textbf{REF}, upper-bound standard error estimates could be produced.
+Using error propagation formulas as in, upper-bound standard error estimates could be produced.
 Figure \@ref(fig:usanx) shows the anxiety ratio metric for 18-29 year-olds  plotted over the survey period.
@@
 $$
-X_{t_i}=\frac{d(s_{j-1},t_i)}{d(s_{j-1},s_{j})}X_{s_j} + \frac{d(t_i,s_{j})}{d(s_{j-1},s_{j})}X_{s_{j-1}} \text{ was used.}
+X_{t_i}=\frac{d(s_{j-1},t_i)}{d(s_{j-1},s_{j})}X_{s_j} + \frac{d(t_i,s_{j})}{d(s_{j-1},s_{j})}X_{s_{j-1}} \text{ was used, where } d \text{ is difference in days.}
 $$
@@
 
-```{r, getgtrends, }
+```{r, getgtrends }
+anx_est_interp <- read.csv(paste0(datadir, "anx_rate_ext.csv"), stringsAsFactors = FALSE)
 ## start and ending time points
-start <- min(as.Date(anxiety$date_end))
-end <- max(as.Date(anxiety$date_end))
+start <- min(as.Date(anx_est_interp$date_end))
+end <- max(as.Date(anx_est_interp$date_end))
 
@@
 
-```{r, gtrendus, fig.cap="Google Trends search popularity of "covid" over time.}
-ggplot(data = covidint, aes(x = as.Date(date), y = hits)) +
+```{r, gtrendus, fig.cap="US Google Trends search popularity of \"covid\" over time."}
+
+ggplot(data = anx_est_interp, aes(x = as.Date(date_end), y = US*100)) +
+    ## geom_line(color = "blue") +
+    geom_line(data = covidint, aes(x = as.Date(date), y = hits)) +
     xlab("") +
     ylab("Search Popularity") +
-    geom_line(alpha = .8)
+    theme_bw()
 ```
 
-## plot the trends together
-ggplot(data = anx_est_interp, aes(x = as.Date(date_end), y = US*100)) +
-    geom_line(color = "blue") +
-    geom_line(data = covidint, aes(x = as.Date(date), y = hits))
+Once again, a generally increasing pattern is noted, with a spike in November which may correspond with the 2020 elections.
+To compare search trends with anxiety data, we will instead find correlation with the first order differences of each data set.
+This will prevent spurious correlation of unit-root processes.
+The correlation was not very significant, even though the the trends do match by eye.
 
+```{r, diffs}
 ## plot differences
@@
 covidint_diffs <- data.frame(date = covidint$date[-length(covidint$date)], 
-                             hitdiffs = log(covidint$hits[-length(covidint$hits)]) - log(covidint$hits[-1]))
+                             hitdiffs = diff(covidint$hits))
 
-ggplot(data = anx_diffs, aes(x = as.Date(date), y = USdiffs)) +
+ggplot(data = anx_diffs, aes(x = as.Date(date), y = 100*USdiffs)) +
     geom_line(color = "blue") +
-    geom_line(data = covidint_diffs, aes(x = as.Date(date), y = hitdiffs))
+    geom_line(data = covidint_diffs, aes(x = as.Date(date), y = hitdiffs)) +
+    xlab("") +
+    ylab("Differences (Anxiety Ratios Scaled)") +
+    theme_bw()
+```
+```{r}
+cor.test(anx_diffs$USdiffs, covidint_diffs$hitdiffs)
+```
 
-Figures:
+We further examined the search trends data across the 50 states.  Correlations calculated between the first order differences in google trends and anxiety data are as follows:
 
-weeks
+```{r, stateweektrends, eval=F}
+## grab trends data for Sat-Sun 7 day periods
+allSun <- as.Date(anx_est_interp$date_end)
+allSat <- as.Date(allSun) - 6
 
-plot of age groups anxiety
+## relative searchs for "covid" by week by state
+weektrends <- vector("list", length(allSun))
+for(i in 1:length(allSun))
+{
+    weektrends[[i]] <- gtrends("lockdown", geo = "US", paste(allSat[i], allSun[i], sep = " "))$interest_by_region
+}
 
-plot of covid?
+## date labels
+names(weektrends) <- allSun
+
+## collapse into dataframe, remove DC, and add date column
+statetrends <- do.call(rbind.data.frame, weektrends)
+statetrends$date_end = substr(rownames(statetrends), 1, 10)
+
+statetrends <- statetrends %>%
+    filter(location != "District of Columbia") %>%
+    mutate(State = sapply(location, function(state) 
+    {
+        state.abb[which(state == state.name)]
+    }))
+
+## write data
+## write.csv(statetrends, paste0(datadir, "statetrends.csv"), row.names = FALSE)
+
+## find first order differences in trends data for each state
+statediffs <- statetrends %>%
+    group_by(State) %>%
+    summarise(hitdiffs = diff(hits), date_end = as.Date(date_end[-1]),
+              .groups = "drop")
+
+## find first order differences in anx data for each state
+anxdiffs <- anx_est_interp %>%
+    pivot_longer(cols = -one_of(c("X", "date_end", "US")), names_to = "State", values_to = "Response") %>%
+    group_by(State) %>%
+    summarise(anxdiffs = diff(Response), date_end = as.Date(date_end[-1]), 
+              .groups = "drop")
+
+## Join dataframes
+diffsdf <- left_join(statediffs, anxdiffs, by = c("State", "date_end")) %>%
+    na.omit()
+
+## write data
+## write.csv(diffsdf, paste0(datadir, "diffsdf_lockdown.csv"), row.names = FALSE)
+```
+
+```{r, covidstates, fig.width=7, fig.height=5, fig.cap="Using the keyword \"covid\"."}
+diffsdf <- read.csv(paste0(datadir, "diffsdf.csv"))
+
+## find correlations on state level
+statecors <- diffsdf %>%
+    group_by(State) %>%
+    summarise(cor = cor(hitdiffs, anxdiffs), .groups = "drop")
+
+## map of US for plotting
+usmapgg <- map_data("state") %>%
+    mutate(region = tools::toTitleCase(region)) %>%
+    mutate(State = state.abb[match(region, state.name)]) %>%
+    left_join(statecors, by = "State")
+
+## states colored by correlation
+ggplot(data = usmapgg) +
+    geom_polygon(aes(long, lat, group = group, fill = cor), color = "grey") +
+    coord_quickmap() +
+    scale_fill_gradient2(low = "blue", high = "red", mid = "white",
+                         midpoint = 0, limit = c(-1,1), space = "Lab",
+                         name="Pearson\nCorrelation") +
+    ggthemes::theme_map()
+```
+
+```{r, lockdown, fig.width=7, fig.height=5, fig.cap="Using the keyword \"lockdown\"."}
+diffsdf <- read.csv(paste0(datadir, "diffsdf_lockdown.csv"))
+
+## find correlations on state level
+statecors <- diffsdf %>%
+    group_by(State) %>%
+    summarise(cor = cor(hitdiffs, anxdiffs), .groups = "drop")
+
+## map of US for plotting
+usmapgg <- map_data("state") %>%
+    mutate(region = tools::toTitleCase(region)) %>%
+    mutate(State = state.abb[match(region, state.name)]) %>%
+    left_join(statecors, by = "State")
+
+## states colored by correlation
+ggplot(data = usmapgg) +
+    geom_polygon(aes(long, lat, group = group, fill = cor), color = "grey") +
+    coord_quickmap() +
+    scale_fill_gradient2(low = "blue", high = "red", mid = "white",
+                         midpoint = 0, limit = c(-1,1), space = "Lab",
+                         name="Pearson\nCorrelation") +
+    ggthemes::theme_map()
+```
+
+As it was seen in the given map, there may be some spatial-temporal trends between the difference in anxiety ratio and differences in Google Search trends for Covid.
+For the sake of this project, the joint significance of correlations of Colorado (where the team is based) and of California was tested using a permutation test.
+Each statistic $T$ was calculated as
+
+$$
+T=corr^2_{\text{California}}+corr^2_{\text{Colorado}}
+$$
+
+The sampling distribution was derived using permutation sampling, wherein each iteration of each state's differences in anxiety ratio is assigned to a randomized state's differences in google trends.
+A sampling distribution was then obtained as $T_{1},\ldots, T_{1000}$.
+
+We obtained $p=0.12$, thus the hypothesis of no significance cannot be rejected.
+Notice that taking into account the nature of data (noisy, interpolation), etc.
+One can argue that with $p-value$ around significance $0.1$ there may be potential for more research in this area.
+
+# Discussion
+
+The anxiety changes in 18–29-year-olds is important for universities, as this age group makes up a large majority of college students.
+However, due to the lack of university public data on student anxiety levels, the finest level of analysis can be done only at the state level.
+
+It is shown that the anxiety ratio of 18–29 year-olds in the United States went through significant perturbations and tended to increase.
+Such changes might negatively affect the performance and achievements of students.
+
+The panel data analysis shows that there are possible links between changes in covid numbers and anxiety ratios.
+Although the direct causal link between variables might be not found, some intermediate variables, like stress about the current situation or the uncertainty about the future, possibly could help to explain such a link.
+
+Additionally, the analysis between changes in google searches and changes in anxiety did not produce a significant relationship but showed to have potential.
+The "eye test" for some clusters of states suggests that some universities can use Google search data to get a general overview of what might be going on in the state and infer about their students.
+
+The methodology described in this report can be easily expanded and should be utilized by universities.
+A possible extension would be to use a network analysis to find a relationship between COVID and anxiety for neighboring states.
+This relationship could allow universities to anticipate increases in anxiety earlier.
+Universities could also collect their own data about student anxiety to establish a more concrete relationship between the statewide anxiety of 18–29-year-olds and the anxiety of their students.
+This analysis could also be applied to more mental health outcomes such as depression.
+
+R code for data collection, cleaning, interpolation where necessary, and analysis are available in the supplemental GitHub repository.
+
+# Bibliography
+
+1.
+Brey, Cristobal de, Thomas D.
+Snyder, Anlan Zhang, and Sally A.
+Dillow.
+“Digest of Education Statistics, 2019.” NCES, February 25, 2021.
+https://nces.ed.gov/ipeds/Search?query=Age&query2=Age&resultType=all&page=1&sortBy=date_desc&collectionYears=2019-20&collectionYears=2018-19&overlayDigestTableId=201035.
+
+2.
+Centers for Disease Control and Prevention, COVID-19 Response.
+COVID-19 Case Surveillance Public Data Access, Summary, and Limitations  https://data.cdc.gov/Case-Surveillance/United-States-COVID-19-Cases-and-Deaths-by-State-o/9mfq-cb36
+
+3. Croissant Y, Millo G (2008). “Panel Data Econometrics in R: The plm Package.” Journal of Statistical Software, 27(2), 1–43. doi: 10.18637/jss.v027.i02.
+
+4.
+Fields JF, Hunter-Childs J, Tersine A, Sisson J, Parker E, Velkoff V, Logan C, and Shin H.
+Design and Operation of the 2020 Household Pulse Survey, 2020.
+U.S.
+Census Bureau https://www.census.gov/programs-surveys/household-pulse-survey/data.html
+
+5.
+Ibrahim, Ahmed K., Shona J.
+Kelly, Clive E.
+Adams, and Cris Glazebrook.
+“A Systematic Review of Studies of Depression Prevalence in University Students.” Journal of Psychiatric Research 47, no.
+3 (March 1, 2013): 391–400. https://doi.org/10.1016/j.jpsychires.2012.11.015.
+
+6. Ku, H.H. “Notes on the Use of Propagation of Error Formulas.” Journal of Research of the National Bureau of Standards: Engineering and Instrumentation 70C, no. 4 (1966): 263–73.
+
+7.
+Lijster, Jasmijn M.
+de, Bram Dierckx, Elisabeth M.W.J.
+Utens, Frank C.
+Verhulst, Carola Zieldorff, Gwen C.
+Dieleman, and Jeroen S.
+Legerstee.
+“The Age of Onset of Anxiety Disorders.” Canadian Journal of Psychiatry.
+Revue Canadienne de Psychiatrie 62, no.
+4 (April 2017): 237–46. https://doi.org/10.1177/0706743716640757.
+
+8. Philippe Massicotte and Dirk Eddelbuettel (2021). gtrendsR: Perform and Display Google Trends Queries. R package version 1.4.8.
+  https://CRAN.R-project.org/package=gtrendsR
+
+9. trends.google.com. "Google Trends." 2021. http://trends.google.com/trends?geo=US

### 2021-04-11 submissiondraftNR.Rmd
final

@@
 Figure \@ref(fig:univpop) shows that ~30\% of 18-29 year-olds in the United States are in higher education and that of those in higher education in the US, ~75\% are between the ages 18 and 29.
-Thus the mental health characteristics of the age bracket 18-29 are will apply to the vast majority of university students.
+Thus the mental health characteristics of the age bracket 18-29 will apply to the vast majority of university students.
 
@@
 
-As described in the technical documentation (LINK), the weeks of the Household Pulse Survey do not represent strict intervals of 7 days.
+As described in the technical documentation, the weeks of the Household Pulse Survey do not represent strict intervals of 7 days.
 For example, "Week 1" refers to data that was collected between April 23rd and May 5th, 2020, (13 days), but Week 9 refers to data that was collected between June 25th and June 30th, 2020 (6 days).
@@
 
-This model allows for state-specific intercepts ($\alpha_s$), autocorrelation in anxiety within a state ($\beta_1$), and an effect of COVID cases on anxiety ($\beta_1$).
+This model allows for state-specific intercepts ($\alpha_s$), autocorrelation in anxiety within a state ($\beta_1$), and an effect of COVID cases on anxiety ($\beta_2$).
 
@@
 We can see that both the lagged anxiety variable and LogCasesRatio are significant predictors.
-The value of the coefficient with $\mbox{COVID}_{s,t}$ shows that if new cases were to double from one week to the next, we expect our response anxiety odds to increase by `r round(fit.states.fixed$coefficients[1] * log(2), digits=4)`.
+The value of the coefficient with $\mbox{COVID}_{s,t}$ shows that if new cases were to double from one week to the next, we expect our response anxiety ratio to increase by `r round(fit.states.fixed$coefficients[1] * log(2), digits=4)`.
 
